### 2010-07-21 trunk/textos/TI 2/texto/01-introduction.tex

\newcommand{\ggp}{ProGReGA}
%\newcommand{\ggp}{PGRGA}
\newcommand{\ggpmeaning}{proportional greedy region growing algorithm}
\newcommand{\ggpk}{\mbox{\ggp-KH}}
\newcommand{\ggpkmeaning}{proportional greedy region growing algorithm keeping heaviest cell}
\newcommand{\ggpf}{\ggp-KF}
\newcommand{\ggpfmeaning}{proportional greedy region growing algorithm keeping usage fraction}
\newcommand{\bfa}{BFBCT}
\newcommand{\bfameaning}{best-fit based cell transference}
\newcommand{\kl}{Kernighan-Lin}
\newcommand{\klalg}{Kernighan-Lin}
\newcommand{\wtodiv}{weight\_to\_divide}
\newcommand{\freecap}{free\_capacity}
\newcommand{\rlist}{region\_list}
\newcommand{\clist}{cell\_list}
\newcommand{\locgroup}{local\_group}
\newcommand{\avguse}{average\_usage}
\newcommand{\wlocal}{local\_weight}
\newcommand{\caplocal}{local\_capacity}
%\newcommand{\wshare}{por\text{\emph{\c{c}}}\tilde{a}o\_da\_carga}
\newcommand{\wshare}{weight\_share}
\newcommand{\wlose}{weight\_to\_lose}
\newcommand{\destreg}{destination\_regions}
\newcommand{\regcap}{free\_capacity}
\newcommand{\wtoreg}{weight\_to\_this\_region}
\newcommand{\misccite}[2]{Available at: #2}
\newcommand{\gamecite}[2]{Available at: #2}

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{times}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}
%[TODO:ajustar TODAS as \refs]
\title{A load balancing scheme for massively multiplayer online games %\thanks{}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Carlos E. B. Bezerra         \and
        Cláudio F. R. Geyer %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Carlos Eduardo Benevides Bezerra, Cláudio Fernando Resin Geyer \at
							Universidade Federal do Rio Grande do Sul \\
              Av. Bento Gonçalves, 9500, Porto Alegre, Brazil \\
              Tel.: +55-51-84065986\\
              Fax.: +55-51-33087308\\
              \email{carlos.bezerra@inf.ufrgs.br, geyer@inf.ufrgs.br}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
%           \and
%           Clï¿½udio Fernando Resin Geyer \at
%           		Universidade Federal do Rio Grande do Sul \\
%              Av. Bento Gonï¿½alves, 9500, Porto Alegre, Brazil \\
%              Tel.: +55-51-99011155\\
%              \email{geyer@inf.ufrgs.br}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
%ORIGINAL ABSTRACT
%MMOGs, or massively multiplayer online games, are applications which require much network bandwidth to function properly. In a distributed MMOG server architecture, with heterogeneous resources, the server nodes may become easily overloaded by the high demand from the players for state updates. Much work has been done in order to distribute the load on the server nodes, but this load is usually defined as the number of players on each server, what is not an ideal measure. Also, the heterogeneity of the system is frequently overlooked. In this work, we propose a balancing scheme which has two main goals: allocate load on the servers proportionally to the each one's power and reduce as much as possible the overhead from the distribution, considering the load as the occupied bandwidth of each server. It is divided in three phases: local selection of servers, balancing and refinement. Four algorithms were proposed, from which \ggp\ is the best for overhead reduction and \ggpf\ is the most suited for reducing player migrations between servers. We also make a review of related works and some comparisons were made, where our approach performed better.
%
%SHRUNK ABSTRACT
In a distributed MMOG (massively multiplayer online game) server architecture, the server nodes may become easily overloaded by the high demand from the players for state updates. Many works propose algorithms to distribute the load on the server nodes, but this load is usually defined as the number of players on each server, what is not an ideal measure. Also, the possible heterogeneity of the system is frequently overlooked. We propose a balancing scheme with two main goals: allocate load on server nodes proportionally to the each one's power and reduce the inter-server communication overhead, considering the load as the occupied bandwidth of each server. Four algorithms were proposed, from which \ggp\ is the best for overhead reduction and \ggpf\ is the most suited for reducing player migrations between servers. We also make a review of related works and some comparisons were made, where our approach performed better.

\keywords{MMOGs \and load balancing \and distributed server \and graph partitioning}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{sec:intro}

@@

When using a client-server architecture, it is necessary that the server intermediates the communication between each pair of players -- assuming that the game is intended to provide guarantees of consistency and resistance to cheating. Obviously, this server will have a large communication load, thus, it must have enough resources (available bandwidth) to meet the demand of the game. We consider here that the main resource to analyse is the available bandwidth, for this is the current bottleneck for MMOGs \cite{feng2007wnn}.

@@

An initial idea might be to distribute the players among servers, so that the number of players on each server would be proportional to that server's bandwidth. However, this distribution would not work, for the burden caused by players also depends on how they are interacting with one another. For example, if the avatars of two players are too distant from each other, probably there will be no interaction between them and therefore the server needs only to update each one of them with the result of their own actions. However, if these avatars are close to each other, each player should be updated not only of his own actions, but also of the actions of the other player.

@@

Differently from other works, we propose here a balancing scheme which considers the upload bandwidth occupation of the server as the load to distribute, what is done among servers with different levels of resources, also reducing the inter-server communication overhead by using a greedy graph partition growing algorithm.

%[TODO:falar das prï¿½xima seï¿½ï¿½es]

\section{Related work}
\label{sec:related}

The servers receive the action performed by a player, calculate its outcome and send it to all interested players, who are usually those whose avatars are close to the avatar of the first player. If two players are split between different servers, each of these need not only to send the state update to the player served by it, but it has also to send the update to the server to which the other player is connected. This server, in turn, forwards that state update to the other player (Figure \ref{fig:overheadgeneration}).

\begin{figure}
 \centering
 \includegraphics[width=0.8\textwidth]{images/overheadgeneration}
 \caption{Overhead caused by the interaction of players connected to different servers}
 \label{fig:overheadgeneration}
\end{figure}

It is perceived, then, that each state will be sent twice for each pair of players who communicate through different servers. This overhead not only causes the waste of resources of the servers, but it also increases the delay to update the status of the replica of the game in the players' machines, damaging the interaction between them.

Therefore, players who are interacting with each other should, ideally, be connected to the same server. However, it is possible that all players are linked through relations of interaction. For example, two avatars of two different players, may be distant from each other, but both could be interacting with a third avatar, between them. Anyway, it is still necessary to divide the players among servers. The question is how many pairs of players and which of them will be divided into different server nodes. It is therefore necessary to decide a criterion to group players. Many works have been done in the past -- such as \cite{devleeschauwer2005dma,lu2006lbm,chen2005lad,duong2003dls,ahmed2008mol} --  trying to find a near optimal, yet fast, load balancing technique.

In \cite{devleeschauwer2005dma}, for example, a few algorithms are proposed. Their idea is to transfer regions managed by overloaded servers to  lightly loaded ones, in a way such that the maximum overload among the servers is reduced. Some very interesting principles are used, as the concepts of microcells and macrocells (which will be described in the next section). However, their model has some differences to ours. For instance, they consider the number of players as the load to distribute, heterogeneous systems are not supported, and the algorithm they use to refine an existing partitioning is simulated annealing. This one may be a little too general, considering that there are algorithms specific to partitioning, such as \cite{kernighan1970ehp} or \cite{fiduccia1982lth}, which may perform better. Finally, some of the algorithms proposed in \cite{devleeschauwer2005dma} do not consider the communication overhead between servers.

It is proposed in \cite{lu2006lbm} a load balancing of cpu usage among the nodes of a cluster. The load balancing is done by an ``off the shelf'' NAT load balancer, which distributes the players' connections among the cluster nodes in a round-robin manner. Thus, each cluster node will end up with the same number of players, on the average. This approach, although it does balance the number of clients per cluster node, communication between the server nodes is not considered, for they are connected via a high-speed and low latency network. Furthermore, in this work we are not concerned with cpu usage, but bandwidth occupation.%OK

Chen et al. propose, in \cite{chen2005lad}, a load shedding technique, in which an overloaded server attempts do shed its load to its neighbors. After finding a lightly loaded neighbor server -- if there is one -- it transfers some of its boundary microcells to that neighbor. To form the group of microcells to transfer, a microcell from the border is chosen, and others are added in a breadth-first search (BFS) order. A very similar work, which also uses BFS, is described in \cite{duong2003dls}. However, these models present a few points which could be improved. For example, a server is considered overloaded when a certain number $N$ of players is reached. As we said before, this is not the best measure, since the load on a server may vary completely, depending on how these $N$ players are distributed across the region managed by the server. Also, the use of BFS to merge two partitions or split one of them may not be ideal. Some other algorithm for graph partitioning applied to distributed systems \cite{kernighan1970ehp,fiduccia1982lth,karypis1999fah,hendrickson1995isg} could be used, for most of these algorithms were designed exactly to balance load and minimize dependence between nodes.

Ahmed and Shirmohammadi propose a microcell oriented load balancing model \cite{ahmed2008mol}. To balance the load, their algorithm, first, finds all clusters of microcells which are managed by the overloaded server. After that, the smallest cluster (in number of microcells) is selected and then, from that cluster, it is chosen the microcell with the lowest communication with other microcells managed by the same server. The chosen microcell is then transferred to the least loaded server. This process is repeated while the server is still overloaded and there are servers which can handle the extra load of the microcells being transferred. The authors define the load on a server, as opposed to other load balancing schemes, as the summing of the message rate of every player that a server must handle, which is a much more realistic measure than the number of players on the server. Also, their model for load balancing tries to minimize the inter-server communication, improving the overall system performance. Their model will be compared to ours, for it takes into consideration many of the aspects that we consider. The comparison will be performed through some simulations, which are described in section \ref{sec:simulations}.

Next, some principles, defined in previous works, will be presented. Based on some of these principles, and on research and implementation carried out, it will be defined the load balancing scheme proposed here.

\subsection{Microcells and macrocells}
\label{sec:micro}

%O balanceamento de carga entre servidores de jogos online maciï¿½amente multijogador ï¿½ fortemente dependente da distribuiï¿½ï¿½o dos avatares dos jogadores atravï¿½s do ambiente virtual. Alï¿½m disso, a depender da concentraï¿½ï¿½o de avatares, pode-se alternar entre uma funï¿½ï¿½o de crescimento de trï¿½fego linear e uma funï¿½ï¿½o quadrï¿½tica. Sendo assim, ï¿½ necessï¿½rio lidar com a localidade dos avatares, de maneira a otimizar o uso de largura de banda do sistema servidor, minimizando, tanto quanto possï¿½vel, o overhead causado pela comunicaï¿½ï¿½o entre jogadores ligados a servidores diferentes.

One way to explore the locality of the players is by grouping them according to the position occupied by their avatars in the virtual environment. %The question would be how to form such groups. One way of doing this would be dividing the world of the game in several cells connected to one another. Each cell would be a part of the world, with its own content and characteristics, which would be assigned to a server node. The easiest way to do this is with a grid of cells of the same size and shape.
%
%[TODO:footnote?]Porï¿½m, o formato e a disposiï¿½ï¿½o destas cï¿½lulas irï¿½ influenciar no trï¿½fego gerado pelas mesmas entre os servidores. Por exemplo, um ambiente bidimensional poderia ser dividido em uma grade de cï¿½lulas quadradas. Neste caso, cada uma destas cï¿½lulas teria oito vizinhos, em mï¿½dia -- cï¿½lulas nas bordas do mapa poderiam ter cinco ou trï¿½s vizinhos apenas. Quanto mais servidores vizinhos, maior o trï¿½fego entre servidores, e maior o overhead causado por esta comunicaï¿½ï¿½o. A distribuiï¿½ï¿½o ideal, entï¿½o, seria utilizando cï¿½lulas hexagonais, cada uma com seis vizinhos. Estudos comprovam que esta ï¿½ a divisï¿½o em cï¿½lulas iguais que permite o menor nï¿½mero de vizinhos por cï¿½lula [TODO:ref]. Outra possibilidade seria utilizando fileiras alternadas de cï¿½lulas quadradas, onde cada fileira seria deslocada o equivalente ï¿½ metade do comprimento de uma cï¿½lula.
%
%[TODO:figura com diferentes tipos de divisï¿½o:sqr/hex/shift]
%
%**	
%An important aspect of this design is that the concept of cell is transparent to the players. They see a wide, single and contiguous world, even if they are repeatedly crossing the boundaries between different cells. Obviously, this requires that the cells communicate, updating and notifying each other of events that occurred near the border between them, and of the migration of players from one to the other.
%
%Although this approach with cells distribute the load among several servers, there are no guarantees that this distribution will be uniform, because of the high mobility of players and the existence of hotspots.
One of the ideas proposed in the literature \cite{devleeschauwer2005dma} follows the principle of dividing the virtual environment in cells of fixed size and position. These cells are relatively small -- or \textbf{microcells} -- and they can be grouped, forming an area called \textbf{macrocell}. Each macrocell is then assigned to a different server, which will manage not a large cell of fixed size and position, but a variable set of small cells. These microcells can then be moved dynamically between different macrocells, maintaining the load on each one of the servers under a tolerable limit.

%[TODO: tirar todas as contraï¿½ï¿½es!]
Obviously, the microcells designated to the same server node do not generate additional traffic to synchronize with each other, but the synchronization overhead of the macrocell is unpredictable, because the number of neighbors of each one of them is not foreknown, for its shape is variable. However, it was demonstrated \cite{devleeschauwer2005dma} that this overhead is compensated by a better distribution of the load between servers in the game. %Figure [TODO: ref] illustrates the division of a virtual environment in two-dimensional microcells.
%It is illustrated in [TODO: ref] in the grouping of these micro macrocï¿½lulas dynamics, which can adapt to the distribution of avatars.

%[TODO: figuraS com microcï¿½lulas e macrocï¿½lulas - fazer vetorial ou copiar? depende do tempo que tiver depois de terminar o texto. divisao em micro e agrupamento em macro]

\subsection{Load balancing in local scope}
\label{sec:local}

In \cite{lee2003sdl}, it is also proposed a scheme for dynamic load balancing for the servers of a multi-server virtual environment. Following the scheme proposed by the authors, an overloaded server starts the process by selecting a number of other servers to be part of the load redistribution. The set of selected servers depends on the load level of the initiating server as well as on the amount of idle resources of the other servers. After the formation of this set, its elements allocate portions of the virtual environment using a graph partitioning algorithm, so that the servers involved have similar final load.

To achieve this goal, the authors also subdivided the virtual environment in rectangular \textbf{cells} -- similar to the microcells --, where the number of servers is much smaller than the number of cells. The cells are grouped into \textbf{regions} -- or macrocells -- and each region is managed by one server (Figure \ref{fig:macromicro}). Each server handles all the interactions between avatars located in the region assigned to it. It receives the inputs of the players controlling these avatars and sends back to them the up-to-date game state.

\begin{figure}
 \centering
 \includegraphics[width=0.4\textwidth]{images/macromicro}
 \caption{Cells grouped into four regions (R$_1$, R$_2$, R$_3$ and R$_4$)}
 \label{fig:macromicro}
\end{figure}

Two cells are called adjacent (or neighbors) if they share a border. Similarly, two regions -- and the servers assigned to them -- are called adjacent (neighbors) if there is a pair of adjacent cells, each of which belonging to one of the two regions. The workload of a cell was defined as the number of avatars present in that cell. The authors assumed that all players receive state updates of the same length and in the same frequency, so that the burden of processing (computing and communication) that a cell requires from a server is proportional to the number of users in that cell. The workload of a region and its designated server is defined as the sum of the individual workloads of the cells which form the region. Each server periodically evaluates its workload and exchange this information with its neighbors. They have assumed, too, that these servers are connected through a high-speed network. Thus, the overhead to exchange workload information among neighbors is limited and considered negligible compared to other costs of the distribution. For the same reason, they also assumed as negligible the overhead of communication between servers in different regions when players are interacting.

The main aspect of the solution proposed by the authors was the use of local information (the server that initiated the balancing process and its neighbors), rather than global information (involving all servers in the balance). When a server is overloaded, it searches for lightly loaded servers close to it in the overlay network. A breadth-first lookup for lightly loaded neighbor servers -- as proposed in \cite{duong2003dls} -- causes a small overhead, but it may not solve the problem efficiently in a few steps, as overloaded servers tend to be adjacent. The global approach, in turn, is able to divide the workload in the most balanced way possible, but its complexity may become too high.

The solution proposed by the authors is then by involving only a subset of servers, such that its cardinality varies according to the need (if the neighbors of the server which triggered the load balancing are also overburdened, more servers are selected). However, this set is formed in a more clever way than by simply using a breadth-first search. The algorithm used to find the servers is described in the next section.

\subsubsection{Selection of a local server group to balance the load}
\label{sec:regselect}

A server starts the balancing when the load assigned to it is beyond its capacity. This server selects a number of other servers to get involved with the distribution. First, it chooses the least loaded server among its neighbors and sends a request that he participates in the load balancing. The chosen server rejects the request if it is already involved in another balancing group, otherwise it responds to the server with the load information of its own neighbors. If the selected neighbor server is unable to absorb all the extra workload of the initiating server, the selection is performed again among the neighbors not only of the overloaded server, but also the neighbors of the already selected servers. The selection continues until the workload of the first server can be absorbed -- that is, the workload of all selected servers becomes smaller than a certain limit.

Figure \ref{fig:lee2} illustrates the operation of the algorithm. All servers have the same capacity, each one being able to handle 100 users. First, the initiator server, $S_6$ is inserted into SELECTED and its neighbors ($S_2, S_5, S_7$ and $S_{10}$) are added to CANDIDATES (Figure \ref{fig:lee2}(a)). So $S_7$, which has the lowest workload among the servers in CANDIDATES, is selected and invited to participate in the load distribution. When $S_7$ sends to $S_6$ the workload information of its neighbors ($S_3, S_6, S_8$ and $S_{11}$), it is inserted into SELECTED and its neighbors, except $S_6$, are added to CANDIDATES (Figure \ref{fig:lee2}(b)). Now, $S_{11}$, which has the lowest workload among servers in CANDIDATES, is selected and invited to participate in the load distribution. However, $S_{11}$ rejects the invitation, because it is involved in another distribution, initiated by $S_{12}$. Thus, $S_{11}$ is removed from CANDIDATES and $S_{10}$ is selected because it now has the lowest workload among all servers in CANDIDATES (Figure \ref{fig:lee2}(c)). This process continues until the average workload is under a pre-defined threshold (Figure \ref{fig:lee2}(d) and Figure \ref{fig:lee2}(e)).

\begin{figure}
 \centering
 \includegraphics[width=1.04\textwidth]{images/localselection}
 \caption{Selecting the group of servers for local rebalance}
 \label{fig:lee2}
\end{figure}

After forming the local server set, the initiating server performs a load rebalancing in such group, using some graph partitioning algorithm. To map the virtual environment to a graph, each cell is represented by a vertex, whose weight equals to the number of avatars in that cell; and every two vertices which represent adjacent cells are connected by an edge. %Though the authors do not provide much detail, the next section will do a brief introduction to graph partitioning algorithms and its use in distributed systems.

%\subsubsection{Using graph partitioning for task distribution}
%\label{sec:regselect}
%
%Um problema clï¿½ssico de alocaï¿½ï¿½o de tarefas em sistemas distribuï¿½dos ï¿½ o da dependï¿½ncia entre elas [TODO:refsDEdistGRAPH]. Tarefas dependentes entre si fazem com que os processadores nos quais elas estï¿½o sendo executadas tenham que se comunicar para que o processamento possa continuar. Isso gera dois problemas principais: em primeiro lugar, o processamento como um todo ï¿½ atrasado por causa do tempo de espera de cada transmissï¿½o e recebimento de mensagens -- supondo que a comunicaï¿½ï¿½o entre os processos seja por meio de mensagens, como ï¿½ o caso do MPI [TODO:ref].
%
%Para resolver este problema, sï¿½o utilizados grafos, da seguinte maneira: o conjunto de tarefas a realizar ï¿½ mapeado em um grafo com pesos. Cada vï¿½rtice representa uma tarefa e cada aresta representa a comunicaï¿½ï¿½o entre as tarefas. O peso de cada vï¿½rtice representa o custo de processamento e o peso da aresta a carga de comunicaï¿½ï¿½o. Para fazer a distribuiï¿½ï¿½o, ï¿½ feito um particionamento do grafo gerado. Cada partiï¿½ï¿½o terï¿½ um conjunto de tarefas que serï¿½o executadas no mesmo nodo e, conseqï¿½entemente, o atraso de comunicaï¿½ï¿½o entre elas serï¿½ relativamente pequeno [TODO:refbli]. No entanto, as arestas do grafo que ligam partiï¿½ï¿½es diferentes representam comunicaï¿½ï¿½o entre diferentes nodos. A Figura [TODO:ref] ilustra essa situaï¿½ï¿½o.
%
%[TODO:figura do grafo de tarefas e dependï¿½ncias, assim como particionamento com corte de arestas e diferentes particionamentos levando a diferentes cortes de aresta]
%
%Os algoritmos de particionamento do grafo de tarefas, geralmente, buscam atingir dois objetivos: gerar partiï¿½ï¿½es de peso aproximadamente igual, ou seja, cada nodo terï¿½ uma carga computacional semelhante ï¿½ dos outros; e minimizar o corte de aresta, ou seja, fazer com que a dependï¿½ncia entre os conjuntos de tarefas seja a menor possï¿½vel, reduzindo a comunicaï¿½ï¿½o entre os nodos de processamento. Por corte de aresta, entende-se o somatï¿½rio dos pesos de todas as arestas que ligam vï¿½rtices que estï¿½o em partiï¿½ï¿½es diferentes. [TODO:reffig]
%
%O problema de particionamento de grafos ï¿½ NP-completo [TODO:ref]. Contudo, jï¿½ foi feita bastante pesquisa nessa ï¿½rea, resultando em heurï¿½sticas que obtï¿½m boas soluï¿½ï¿½es, muitas vezes se aproximando de valores ï¿½timos [TODO:ref]. Uma das heurï¿½sticas mais conhecidas ï¿½ a de Kernighan e Lin \cite{kernighan1970ehp}, que tem como objetivo, partindo de duas partiï¿½ï¿½es iniciais quaisquer, chegar a duas partiï¿½ï¿½es com peso semelhante e corte de aresta reduzido. O algoritmo de Fiduccia e Mattheyses \cite{fiduccia1982lth} incrementa a soluï¿½ï¿½o de Kernighan e Lin, generalizando-a para hipergrafos, alï¿½m de permitir que cada partiï¿½ï¿½o tenha uma fraï¿½ï¿½o diferente do peso total do hipergrafo.
%
%Existem outros algoritmos mais recentes, como os baseados em divisï¿½o espectral [TODO:ref-33-e-20-karypisirreggraphs]. Este ï¿½ computacionalmente caro, por envolver complexos de ï¿½lgebra linear, precisando autovetor associado ao menor autovalor da matrix laplaciana associada ao grafo [TODO:ref]. No entanto, o particionamento resultante ï¿½ considerado excelente para uma vasta gama de problemas [TODO:ref]. Existe tambï¿½m a abordagem em vï¿½rios nï¿½veis [TODO:citemultilevel], atravï¿½s da simplifaï¿½ï¿½o do grafo por meio de contraï¿½ï¿½o dos mesmos e sobreposiï¿½ï¿½o das arestas correspondentes. Com o grafo mais simples, ï¿½ aplicado algum algoritmo de particionamente e entï¿½o ele se desdobra, jï¿½ particionado, atï¿½ o grafo original. Eventualmente, refinamentos no particionamento podem ser necessï¿½rios. Esta ï¿½ltima tï¿½cnica ï¿½ especialmente ï¿½til em grafos com grande nï¿½mero de vï¿½rtices, na ordem de centenas a milhares, ou mais. Na figura [TODO:ref] ï¿½ ilustrado esse particionamento em vï¿½rios nï¿½veis.
%
%[TODO:figura do multilevel]
%
%Contudo, MMOGs sï¿½o geralmente aplicaï¿½ï¿½es de tempo-real, alï¿½m de a carga imposta sobre os servidores mudar constantemente em tempo de execuï¿½ï¿½o. Por esse motivo, o algoritmo de particionamento utilizado deve ser rï¿½pido, tornando inviï¿½vel o uso da divisï¿½o espectral. ï¿½ proposto neste trabalho utilizar um algoritmo guloso, mais simples, definido em [TODO:aOUTRArefdekarypis], que pode nï¿½o atingir soluï¿½ï¿½es tï¿½o boas quanto as da divisï¿½o espectral, porï¿½m ï¿½ consideravelmente mais rï¿½pido. Alï¿½m disso, o custo para encontrar uma soluï¿½ï¿½o muito prï¿½xima do ï¿½timo nï¿½o se justifica no contexto de MMOGs, pelo fato dos pesos mudarem constantemente, precisando de um novo particionamento em relativamente pouco tempo. Jï¿½ quanto ao particionamento em vï¿½rios nï¿½veis, seu uso nï¿½o foi considerado necessï¿½rio pois o nï¿½mero de vï¿½rtices do grafo que representarï¿½ o ambiente virtual ï¿½ relativamente pequeno.
%
%%TODO:ver esse "definiï¿½ï¿½es e parametros"
%Nas prï¿½ximas seï¿½ï¿½es, serï¿½ apresentada a soluï¿½ï¿½o proposta neste trabalho para balanceamento de carga, comeï¿½ando pelas definiï¿½ï¿½es e parï¿½metros utilizados, assim como serï¿½ descrito como o problema foi mapeado para grafos e quais os objetivos e critï¿½rios do esquema proposto.

\section{Proposed load balancing scheme}
\label{sec:scheme}

In the previous section it was presented some existing works regarding load balancing in MMOGs when using multiple servers to provide the support network. The load balancing scheme proposed in this work is based on some of the principles in the literature. One of them is the division of the virtual environment in microcells, for later grouping in macrocells. This is a relatively simple way of addressing the issue of the avatars' movement dynamics through the game world, by transferring the microcells dynamically according to need.

It will also be used the idea of balancing based only on local information -- each server, when needing to reduce its workload, selects only a few other servers to join a local load rebalancing. Thus, we can greatly reduce the complexity of balancing because it will not be necessary that all servers in the game exchange messages among themselves every time that any one of them is overloaded.

To define our load balancing approach, we cannot ignore that, usually, the traffic generated by the players is not simply linear, but square for each cluster of players. Such misunderstanding can generate considerable differences between the actual load of each server and the load estimated by the balancing algorithm. Another point to consider is the overhead, both in the delay of sending messages, as in the use of bandwidth of the servers, when players connected to different servers are interacting with each other. Some works assume that the servers are all in the same high-speed and low latency local-area network, and therefore overhead is negligible. However, when considering a geographically distributed server system, this assumption cannot be made. This overhead must be taken into account, no matter which load balancing algorithm is being used.

Another important point is that the main criterion we use when balancing the load of MMOG servers is the bandwidth, and not processing power. Several games, such as Age of Empires \cite{ageofempires}, include simulations of virtual environments with hundreds or thousands of entities, which are performed smoothly on current personal computers. However, if this game was multiplayer and each of these entities was controlled by a different player on the Internet, it would most likely generate a traffic amount which would hardly be supported by a domestic connection \cite{feng2007wnn}.

%TODO: explicar porque cresce linearmente
Moreover, the upload bandwidth must be taken into account, much more than download. This happens for two reasons: first, the usage of the download bandwidth of each server node grows linearly with the number of players connected to it, while the usage of the upload bandwidth may have a quadratic growth, getting quickly overwhelmed; second, domestic connections -- which are majority in volunteer peer-to-peer systems -- usually have an upload bandwidth much lower than the download one.

There are also other issues which must not be overlooked. One of them is that the server system is probably not homogenous -- considering that it is geographically distributed with, likely, different connections to the Internet. Therefore, one cannot assume that the servers have the same amount of cpu power or network bandwidth. % Another problem, specific to the algorithm proposed by \cite{lee2003sdl}, is that the criterion to allocate to each server a load equivalent to 90\% of its capacity in order to avoid constant rebalancing is weak. There are no guarantees that the server system has 11.11\% more total capacity than the necessary. On the contrary, it should adapt to situations of widespread overload. Moreover, the criterion for stopping is weak because the algorithm does not stop when the whole system is loaded above 90\% of its capacity, or if the overloaded server has no neighbors capable of receiving more load.
In the next section, we make some definitions that will be necessary to specify our load balancing scheme.

%[TODO:introduzir as seï¿½ï¿½es]

%However, an important idea that was suggested was the use of graphs to represent the virtual environment, and the use of graph partitioning algorithms to perform load balancing. In the following section, will be a brief introduction about this principle, which is the basis of load balancing scheme proposed here.

\subsection{Definitions}
\label{sec:def}

We also use the idea of mapping the virtual environment on a graph. The graph will then be partitioned to distribute the workload of the game between the different servers. It is necessary first to define some terms which will be used on the proposed algorithms.

%TODO:definir um por um em detalhes, depois enumerï¿½-los no parï¿½grafo acima ^ para o leitor nï¿½o se perder

\begin{itemize}
	\item \textbf{Server}: here, server is defined as a node belonging to the distributed system to serve the game. Each server can be assigned a single region;
	\item \textbf{Server power}: the server power, $p(S)$ is a numerical value proportional to the server's upload bandwidth;
	\item \textbf{Server power fraction}: given a set of servers $Servers$ = $\{S_1, S_2, ..., S_n\}$, the power fraction of a server $S$, $frac_p(S)$, is equal to its power divided by the summed power of all servers in $Servers$:

	%**
		\begin{center}
			$frac_p(S) = \frac{\displaystyle p(S)}{\displaystyle\sum_{i=1}^{n} p(S_i)}$
		\end{center}

	\item \textbf{System power}: the total power of the system, $P_{total}$, equals the sum of the powers of the $n$ servers which form it:

		\begin{center}
			$\displaystyle P_{total} = \sum_{i=1}^{n} p(Si)$
		\end{center}

	\item \textbf{Cell}: similar to the microcells, it is considered here the environment being divided into small cells, each one with fixed size and position. If two cells share a border, they are said to be \textbf{adjacent} or \textbf{neighbors};
	\item \textbf{Region}: the cells are grouped, forming what is called regions. Usually these areas are contiguous, although in some cases the subgraph that represents them may be disconnected, resulting in the presence of cells isolated from each other. Each region is assigned to a server, and only one, and $s(R)$ is the server associated to the region $R$. It may be referred throughout the text to region's ``power'', which in fact refers to the power of the server associated with that region, i.e. $p(s(R))$;
	%[TODO:ver se realmente precisa do item interaï¿½ï¿½o, ou se nao eh melhor usar relevancia mesmo]
	\item \textbf{Relevance}: the relevance of an avatar $A_j$ to another one, $A_i$, determines the frequency of updates of the state of $A_j$ the server should send to the player controlling $A_i$ \cite{bezerra2008a3}. It may be represented by the function $R(A_i,A_j)$;
	\item \textbf{Avatar's weight}: to each avatar there are various other entities (here, we only consider avatars) of the game, each with a frequency of state updates that need to be sent to the player who controls that avatar. Thus, for each avatar $A$ , its individual weight -- or of upload bandwidth that the server uses to send state updates to its player -- $w_a(A)$ depends on which other entities are relevant to it, and how much. 
	%**
	Let $\{A_1, A_2, ..., A_t\}$ be the set of all avatars in the virtual environment, we have:

		\begin{center}
			$\displaystyle w_a(A) = \sum_{i=1}^{t} R(A,Ai)$
		\end{center}

	\item \textbf{Cell's weight}: here, the total weight of a cell (or the use of upload bandwidth of its server) will be equal to the sum of the individual weights of the avatars in it. Consider cell $C$, where the avatars $\{A_1, A_2, ..., A_n\}$ are present. Also, consider that the avatars $\{A_1, A_2, ..., A_t\}$ are all the avatars in the virtual environment. The weight of this cell, $w_c(C)$, is:

		\begin{center}		
			$\displaystyle w_c(C) = \sum_{i=1}^{n} w(Ai) = \sum_{i=1}^{n} \sum_{j=1}^{t} R(A_i,A_j)$
		\end{center}

	\item \textbf{Region's weight}: the weight of a region is the sum of the weights of the cells that compose it. Let $R$ be the region formed by $\{C_1, C_2, ..., C_p\}$. The weight of $R$ is:

		\begin{center}
			$\displaystyle w_r(R) = \sum_{i=1}^{p} w_c(Ci)$
		\end{center}

		\item \textbf{Region's weight fraction}: given a set of regions $Regions  = \{R_1, R_2, ..., R_n\}$, the weight fraction of $R$, relative to $Regions$, is:

>>>>

% aqui comeca o texto propriamente dito

% introducao
\chapter{Introduction}

@@

When using a client-server architecture, it is necessary that the server intermediates the communication between each pair of players -- assuming that the game is intended to provide guarantees of consistency and resistance to cheating. Obviously, this server will have a large communication load, thus, it must have enough resources (available bandwidth) to meet the demand of the game. For this reason, it must be considered that the main resource to analyse is the available bandwidth, for this is the current bottleneck for MMOGs \cite{feng2007wnn}.

@@

An initial idea might be to distribute the players among servers, so that the number of players on each server would be proportional to that server's bandwidth. However, this distribution would not work, for the burden caused by players also depends on how they are interacting with one another. For example, if the avatars of two players are too distant from each other, probably there will be no interaction between them and therefore the server needs only to update each one of them with the result of their own actions. However, if these avatars are close to each other, each player should be updated not only about the results of his own actions, but also of the actions of the other player.

@@

In this work, it is presented a few recent papers which try to solve this problem, achieving a significant fair load distribution. Besides reviewing these works, some comments are added based on the advantages and weaknesses of each solution presented by the authors. In the end of this paper, a new idea is presented, suggesting the use of a generic binary space partition (BSP) tree.
